Deep Probabilistic Models for Camera Geo-Calibration

### 2018-08-14 chapters/introduction/introduction.tex
minor changes

\subsection{Geo-Calibration}
\todo{add deterministic methods}

Automatic image geo-calibration continues to grow in importance as a
direct result of the increasing amount of imagery available
via the Internet. Conceptually geo-calibration task is
straightforward: given an image, identify the time and the location it
was captured in the world and the direction the camera was facing.
Solving this problem is of great value for a wide variety of fields,
with potential applications ranging from the forensic
sciences~\cite{stylianou13jane} to crowd-sourced environmental
monitoring~\cite{zhang2012mining}.

@@

However, recognizing the geo-location and geo-orientation of an
arbitrary outdoor image is an extremely challenging task.  Many
methods have been proposed; the most common approach is to build a
large database of images with known location and localize a query
image using either local~\cite{li2010location,schindler2008detecting}
or global~\cite{hays2008im2gps,doersch2012what} image features.  This
approach is not applicable when no nearby ground-level imagery exists
in the reference database, such as when the image was not captured
near a popular tourist destination.  Even when reference imagery is
available, the appearance of the objects may not be visually
distinctive, for example a train track or a body of water.

\subsection{Deep Learning in Geo-Calibration}
In recent years, deep neural networks (DNNs) were proven to be
extremely successful in many computer vision areas (\todo{citations}). 
The deep neural networks prove to be a good tool to extract high-level
image representations, this valuable ability offers great a help for
researchers to conquer the problems like geo-localization and camera
calibration.
Weyand and Kostrikov \etal~\cite{planet} propose a convolutional
network to predict the geographic location given an input
image.~\cite{walch2017image} applies LSTM on CNN features learned for
camera localization.

@@

\subsection{Ground-to-Aerial Geo-Calibration}
Data-driven scene matching approaches are widely used to localize a
query image by registering in a large dataset of geo-tagged
images~\cite{im2gps, li2010location,zamir2010accurate}.
Benefited from the fast growing monitoring satellite and drone markets,
satellite/aerial imagery becomes more and more available than ever
before, which makes it a rich source for
reference database of geo-tagged images. We refer to the
geo-calibration methods which uses the aerial imagery as geographic
reference as {\em Ground-to-Aerial Geo-Calibration}.
%
Cross-view image pair is the basic element for ground-to-aerial
geo-calibration. It consists of a ground-level image and an overhead
image at the same location. In some cases, two images are also aligned
in orientation. 
%
Recent work on cross-view image
geo-localization~\cite{lin2013cross,lin2015learning,workman2015geocnn,workman2015wide}
has shown that convolutional neural networks are capable of extracting
features from geo-tagged aerial imagery that can be matched to features extracted
from the inquiry ground imagery.  Vo \etal~\cite{vo2016localizing} extend this
line of work, demonstrating improved geo-localization performance by
applying an auxiliary loss function to regress the ground-level camera
orientation with respect to the aerial image. Hu and Feng
\etal~\cite{mh2018cvm} achieve the state-of-the-art performance in
geo-localization by extracting cross-domain global features using
NetVLAD~\cite{arandjelovic2016netvlad} technique.

@@

Automatic feature engineering has drawn a lot of attention over the
past years in the computer vision community due to its efficiency and
not requiring expert knowledge about the data. Deep neural networks
are wildly accepted by researchers as good tools to learn high-level
image representations.
The most commonly used approach for deep image feature learning is
full supervision, which usually requires manually annotated
data~\cite{yosinski2014transferable,zhou2016learning,wen2016discriminative}.
Though full supervision work well in many applications, there are
exponentially more image resources without manual annotations (cell
phone imagery, YouTube videos\etc), which show a huge potential for
visual feature learning.  In recent years, people start studying
various approaches to exploit those unlabeled data. 
%
Recent work has explored the use of self-supervision (sometimes
referred to as unsupervised learning or pretext tasks) for training
deep neural networks that capture useful visual
representations~\cite{doersch2015unsupervised,pathak2016context}. 
These methods typically
exploit some known quantity of the data (e.g., pixel color values) to
avoid expensive manual annotation. As a byproduct, a useful visual
representation is learned.
For example, Zhang et al.~\cite{zhang2016colorful} show how image
colorization (synthesizing colors for a grayscale image) is a powerful
pretext task for learning visual representations. Pathak et
al.~\cite{pathak2017learning} exploit low-level motion-based grouping
cues for unsupervised feature learning.  
%
As new learning technique address the lack of massive amounts of
labeled data, domain adaptation adapt the feature generating model to
another feature domain so that it improve the performance when dealing
with new input
data~\cite{fernando2013unsupervised,fernando2015joint,saenko2010adapting,wang2016actions,tinghui2016flow}.
Duan \etal~\cite{duan2012learning} proposed to learn a linear
projection to transfer from the source feature subspace to the target
subspace. Sohn \etal~\cite{sohn2017unsupervised} improved the
performance of video face recognition using an adversarial-based
approach to adapt the network to unlabled video imagery.

@@

\section{Our Approach}
Our approach consists of two big steps. First, we propose a general
probabilistic model to jointly estimate the geo-calibration
parameters. The key for the success of this model is to construct good
constraint functions which describe the fitness between the calibration
parameters and the images. As the second step of our approach, we
dismantle the full geo-calibration into several partial
geo-calibration tasks, each of which could provides a strong constraint
function that fit in the general model. The architecture of
our approaches refers to \figref{intro:architecture}.
%
In the end, we show that learning to geo-calibrate a camera is helpful
for high-level image feature extracting.

@@

  \item \textbf{Constraint for Time and Coarse Geographic Locations:}
  In this work, we demonstrate to estimate the capture time and the
  camera geographic location with the input image.
  Our network has two branches to predict the image capture time and
  the geographic location jointly.
  For time estimation, our network is able to predict the
  distribution over the month of the year and the global hour of the
  day when the input image has been captured.
  For camera geo-localization, We provide two approaches to obtain the
  probability distribution over the camera geographic location. Most
  straightforwardly, our network estimate the categorical distribution
  over world into a $N \times M$ bins. Better results can be procured
  with temporal inputs (image timestamps).  Our algorithm can also
  compute the probability at arbitrary geographic location, with lower
  accuracy than the straightforward method.
  We present the algorithm details in \chapref{whenwhere}.
  \newline

  \item \textbf{Constraint for Fine Geographic Locations and Orientations:}
  In our previous method, we present that our algorithm can estimate
  the camera location distribution in low-res world division. In this
  work we focus on refining the geo-localization results with the help
  of overhead imagery and on estimating the camera geo-orientation. 
  We first learn to segment the ground scene from its paired aerial image.
  When refining the camera geographic location, we compare the
  semantic layout of the ground image with those transferred from
  overhead images sampled from different locations and orientations. The
  optimal prediction is found when the semantic layouts from these
  two sources match best. For probabilistic outputs, we assign
  a score to each location and orientation setting,
  based on how well the semantic layouts match to each other.
  We present the algorithm details in \chapref{crosstransf}.
  \newline

@@

\subsection{Geo-Calibration for Feature Learning}
Our methods focus on lavaging the camera geographic information to
deal with the challenges of lacking manually annotated data during DNN
training. The insight is that the camera geographic information
encodes semantic information because it is related to the
appearances of the scenes.

\begin{itemize}[noitemsep]
  \item \textbf{Ground-to-Aerial Feature Learning:}
  Current machine learning methods address problems that involves mostly
  the ground-level imagery. Tons of dataset targeting on ground imagery are
  annotated while much fewer are done for the overhead imagery. In
  \chapref{crosstransf}, we proposed a method to learn aerial image
  semantic segmentation with ground-level ground truth. Our method
  exploits the spatial correspondences between images in cross-view
  pairs. 
  In this work, we try to transfer the semantics from the ground
  imagery to the overhead imagery by identifying the latent geometry
  correspondences between these two.  Similar to the monocular depth
  estimation methods that driven by the technique called {\em
  Reprojection Losses}~\cite{garg2016unsupervised,
  godard2017unsupervised,zhou2017unsupervised, yan2016perspective}, our
  network learns to geometrically project pixels from the overhead
  images to the ground images, and to segment the overhead images
  jointly.
  \newline

  \item \textbf{Geo-Temporal Feature Learning:}
  In \chapref{whenwhere}, we demonstrate that our algorithm is able to
  learn geo-temporal image representations with geographic locations and
  capture time.
  Our work is most similar to Li and Snavely~\cite{li2018learning},
  who propose an approach for learning intrinsic images by observing
  image sequences with varying illumination over time. While their
  targets are reflectance and shading images of the scene, we consider
  two novel pretext tasks, time and location estimation.
  By learning to predict its geographic location and capture time from the
  query image, our network learns to extract geo-temporal features
  from the image.

>>>>

\subsection{Geo-Calibration}
%\todo{add deterministic methods}
%
Automatic image geo-calibration continues to grow in importance as a
direct result of the increasing amount of imagery available
via the Internet. Conceptually geo-calibration task is
straightforward: given an image, identify the time and the location it
was captured in the world and the direction the camera was facing.
Solving this problem is of great value for a wide variety of fields,
with potential applications ranging from the forensic
sciences~\cite{stylianou13jane} to crowd-sourced environmental
monitoring~\cite{zhang2012mining}.

@@

However, recognizing the geo-location and geo-orientation of an
arbitrary outdoor image is an extremely challenging task.  Many
methods have been proposed; the most common approach is to build a
large database of images with known location and localize a query
image using either local~\cite{li2010location,schindler2008detecting}
or global~\cite{hays2008im2gps,doersch2012what} image features. 
%This approach is not applicable when no nearby ground-level imagery
%exists in the reference database, such as when the image was not
%captured near a popular tourist destination.  Even when reference
%imagery is available, the appearance of the objects may not be
%visually distinctive, for example a train track or a body of water.

\subsection{Deep Learning in Geo-Calibration}
In recent years, deep neural networks (DNNs) were proven to be
extremely successful in many computer vision areas (\todo{citations}). 
The deep neural networks prove to be a good tool to extract high-level
image representations, this valuable ability offers great a help for
researchers to conquer the problems like geo-localization and camera
calibration.
Weyand and Kostrikov \etal~\cite{planet} propose a convolutional
network to predict the geographic location of the input
image.~\cite{walch2017image} applies LSTM on CNN features learned for
camera localization.

@@

\subsection{Ground-to-Aerial Geo-Calibration}
Data-driven scene matching approaches are widely used to localize a
query image by registering in a large dataset of geo-tagged
images~\cite{im2gps, li2010location,zamir2010accurate}.
Benefited from the fast growing monitoring satellite and drone markets,
geo-tagged satellite/aerial imagery becomes more and more available
than ever before, which makes it a rich source for building
reference database. We refer to the geo-calibration methods which use
the aerial imagery as geographic reference as {\em Ground-to-Aerial
Geo-Calibration}.

Cross-view image pair is the basic element for ground-to-aerial
geo-calibration. It consists of a ground-level image and an overhead
image at the same location. In some cases, two images may also be aligned
in orientation.

Recent work on ground-to-aerial
geo-localization~\cite{lin2013cross,lin2015learning,workman2015geocnn,workman2015wide}
has shown that convolutional neural networks are capable of extracting
features from geo-tagged aerial imagery that can be matched to features extracted
from the inquiry ground imagery.  Vo \etal~\cite{vo2016localizing} extend this
line of work, demonstrating improved geo-localization performance by
applying an auxiliary loss function to regress the ground-level camera
orientation with respect to the aerial image. Hu and Feng
\etal~\cite{mh2018cvm} achieve the state-of-the-art performance in
geo-localization by extracting cross-domain global features using
the NetVLAD~\cite{arandjelovic2016netvlad} technique.

@@

Automatic feature engineering has drawn a lot of attention over the
past years in the computer vision community due to its efficiency and
not requiring expert knowledge about the data. Deep neural networks
are wildly accepted by researchers as good tools to learn high-level
image representations.
The most commonly used approach for deep feature learning is
full supervision, which usually requires manually annotated
data~\cite{yosinski2014transferable,zhou2016learning,wen2016discriminative}.
Though full supervision work well in many applications, there are
exponentially more image resources without manual annotation (cell
phone imagery, YouTube videos\etc), which show a huge potential for
visual feature learning.

To exploit these unlabeled data, recent work has explored the use of
self-supervision (sometimes referred to as unsupervised learning or
pretext tasks) for training deep neural networks that capture useful
visual representations~\cite{doersch2015unsupervised,pathak2016context}. 
These methods typically
exploit some known quantity of the data (e.g., pixel color values) to
avoid expensive manual annotation. As a byproduct, a useful visual
representation is learned.
For example, Zhang et al.~\cite{zhang2016colorful} show how image
colorization (synthesizing colors for a grayscale image) is a powerful
pretext task for learning visual representations. Pathak et
al.~\cite{pathak2017learning} exploit low-level motion-based grouping
cues for unsupervised feature learning.  
%
As new learning technique address the lack of massive amounts of
labeled data, domain adaptation adapt the feature generating model to
another feature domain so that it improve the performance when dealing
with new input
data~\cite{fernando2013unsupervised,fernando2015joint,saenko2010adapting,wang2016actions,tinghui2016flow}.
Duan \etal~\cite{duan2012learning} proposed to learn a linear
projection to transfer from the source feature subspace to the target
subspace. Sohn \etal~\cite{sohn2017unsupervised} improved the
performance of video face recognition using an adversarial-based
approach to adapt the network to unlabled video imagery.

@@

\section{Our Approach}
Our approach consists of two big steps. First, we propose a general
probabilistic model to jointly estimate the geo-calibration
parameters. The key for the success of this model is to construct good
constraint functions which describe the fitness between the calibration
parameters and the images. As the second step of our approach, we
dismantle the full geo-calibration into several partial
geo-calibration tasks, each of which could provides a strong constraint
function that fits in the general model. The architecture of
our approaches refers to \figref{intro:architecture}.
%
In the end, we show that learning to geo-calibrate a camera is helpful
for high-level image feature extracting.

@@

  \item \textbf{Constraint for Time and Coarse Geographic Locations:}
  In this work, we demonstrate to estimate the capture time and the
  camera geographic location given the input image.
  Our network has two heads to predict the image
  capture time and the geographic location jointly.
  For time estimation, our network is able to predict the
  joint distribution over the month of the year and the global hour of the
  day when the input image has been captured.
  For camera geo-localization, We provide two approaches to obtain the
  probability distribution over the camera geographic location. Most
  straightforwardly, our network estimate the categorical distribution
  over location of the world, which is represented as a $N \times M$
  meshgrid. Better results can be procured with image capture time as
  auxiliary input (if it is known). 
  %Our algorithm can also compute the probability at arbitrary
  %geographic location, with lower accuracy than the straightforward
  %method.
  We will present the algorithm details in \chapref{whenwhere}.
  \newline

  \item \textbf{Constraint for Fine Geographic Locations and Orientations:}
  In our previous method, we present that our algorithm can estimate
  the camera location distribution in low-res world division. In this
  work we focus on refining the geo-localization results with the help
  of overhead imagery and on estimating geo-orientation of the camera. 
  We first learn to segment the ground scene from its paired aerial image.
  When refining the camera geographic location, we compare the
  semantic layout of the ground image with the semantic layout
  transferred from overhead images, which are sampled in different
  locations and orientation angles. The optimal prediction is found when the
  semantic layouts from these two sources match best. For probabilistic
  outputs, we assign a score to each location and orientation setting,
  based on how well the semantic layouts match to each other. We will
  present the algorithm details in \chapref{crosstransf}.
  \newline

@@

\subsection{Geo-Calibration for Feature Learning}
Our methods also focus on lavaging the camera geographic information to
deal with the challenges of lacking manually annotated data during DNN
training. The insight is that the camera geographic information
encodes semantic information because it is related to the
appearances of the scenes.

\begin{itemize}[noitemsep]
  \item \textbf{Ground-to-Aerial Feature Learning:}
  Current machine learning methods address problems that involves mostly
  the ground-level imagery. Tons of dataset targeting on ground imagery are
  annotated while much fewer are done for the satellite/aerial imagery. In
  \chapref{crosstransf}, we proposed a method to learn the aerial image
  semantic segmentation with only ground-level ground truth. Our method
  exploits the spatial correspondences between images in cross-view
  pairs. 
  In this work, we try to transfer the semantics from the ground
  imagery to the aerial imagery by identifying the latent geometry
  correspondences between these two. Similar to the monocular depth
  estimation methods that driven by the technique called {\em
  Reprojection Losses}~\cite{garg2016unsupervised,
  godard2017unsupervised,zhou2017unsupervised, yan2016perspective}, our
  network learns to geometrically project pixels from the aerial
  images to the ground images, and to segment the aerial images
  jointly.
  \newline

  \item \textbf{Geo-Temporal Feature Learning:}
  In \chapref{whenwhere}, we demonstrate that our algorithm is able to
  learn geo-temporal image representations with geographic locations and
  capture time.
  Our work is most similar to Li and Snavely~\cite{li2018learning},
  who propose an approach for learning intrinsic images by observing
  image sequences with varying illumination over time. While their
  targets are reflectance and shading images of the scene, we consider
  two novel pretext tasks, time and location estimation.
  By learning to predict the geographic location and capture time of
  the query image, our network learns to extract geo-temporal features
  from the image.
